---
title: "NYC Restaurant Inspection Data in Plotly"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Let's take a look at data from NYC restaurant inspections. This data set contains information about the results of food safety inspections conducted by the New York City Department of Health and Mental Hygiene. It was taken from the DOHMH via NYC Open Data in October 2017. To learn more, please visit the data set's [host website](https://p8105.com/dataset_restaurant_inspections.html), where additional links to the current raw data and a data dictionary can also be found. 

### Import and inspect the data

```{r inspect, message = FALSE}
data("rest_inspec")

rest_inspec %>% 
  filter(boro == "MANHATTAN") %>% 
  select(cuisine_description) %>% 
  group_by(cuisine_description) %>% 
  summarize(fqcy = n()) %>% 
  filter(fqcy >= 2000) %>% 
  knitr::kable(
    col.names = c("Cuisine", "Frequency")
  )
```

Due to the size of the dataset (`r nrow(rest_inspec)` observations), let's just look at establishments for cuisines that have over 2000 restaurants in Manhattan. We will also clean up some of the names of cuisines, and drop unnecessary variables for our analysis.

```{r tidy, message = FALSE}
rest_list = 
  rest_inspec %>% 
  filter(boro == "MANHATTAN") %>% 
  select(cuisine_description) %>% 
  group_by(cuisine_description) %>% 
  summarize(fqcy = n()) %>% 
  filter(fqcy >= 2000) %>% 
  select(-fqcy)

restaurants_df = 
  rest_inspec %>% 
  inner_join(rest_list, by = "cuisine_description") %>% 
  filter(boro == "MANHATTAN") %>% 
  mutate(
    cuisine_description = recode(cuisine_description, 
           "Café/Coffee/Tea" = "Cafe",
           "Latin (Cuban, Dominican, Puerto Rican, South & Central American)" = "Latin")
  ) %>% 
  select(-c(camis, building, street, phone, record_date)) %>% 
  drop_na(score, grade, critical_flag, inspection_type)
```

### Pizza Plot

First, an essential cuisine in New York: pizza.

```{r pizza, warning = FALSE}
pizza_df = 
  restaurants_df %>% 
  filter(
    cuisine_description == c("Pizza", "Pizza/Italian")
  )
```

Let's examine the distribution of inspection score (higher score means worse performance) and grade of pizza restaurant, separated by severity of violation (critical or not). 

```{r pizza_plot, warning = FALSE}
pizza_plot = 
  pizza_df %>% 
  filter(
    grade %in% c("A", "B", "C"),
    critical_flag != "Not Applicable"
  ) %>% 
  mutate(
    grade = as.factor(grade),
    text_label = str_c("Grade: ", grade, "\nScore: ", score)     
  ) %>% 
  plot_ly(x = ~grade, y = ~score, color = ~critical_flag, text = ~text_label,
    type = "box", colors = "viridis", alpha = 0.5
  ) %>% 
  layout(boxmode = "group", boxgap = .25, boxgroupgap = 0.5)

pizza_plot
```

As we expected, a better grade is associated with a score on inspection. There do appear to be a few outliers in the "C" rated category, with a few low severity scores (under 10) still resulted in a poor health grade, and one critical violation with a total 60 points (don't eat there). 

### Restaurant inspection performance by cuisine

Now, for all our selected cuisines, let's investigate which are routinely scored the best, and which often have poor ratings.

```{r cuisine_plot, message = FALSE}
cuisine_plot = 
  restaurants_df %>% 
  filter(
    grade %in% c("A", "B", "C"),
    critical_flag != "Not Applicable"
  ) %>% 
  group_by(cuisine_description, grade) %>% 
  summarize(med_score = median(score)) %>% 
  mutate(
    text_label = str_c(
      "", cuisine_description,
      "\nMedian Score: ", med_score, 
      "\nGrade: ", grade
      )
  ) %>% 
  plot_ly(
    x = ~cuisine_description, y = ~med_score, color = ~grade,
    text = ~text_label, type = "bar", colors = "viridis"
  ) %>% 
  layout(xaxis = list(title = "", tickangle = -45), 
         yaxis = list(title = "Median Score"),
         bargap = 0.25,
         legend = list(title = list(text = "Grade")))

cuisine_plot
```

There doesn't seem to be much variation in median score across A and B ratings; they seem to fall around 10 and 20 points for all cuisines, respectively. However, we do see more variation in median score for C grades. Spanish restaurants tend to have higher scores (again, meaning worse ratings) with a median of 41, the highest for the group; on the other end, sandwich shops tend to have lower median scores at 28 for a C grade. 

### Restaurant Closures

Finally, going back to our full dataset, let's take a look at restaurants who were closed by DOHMH for critical violations, across all 5 boroughs, within 2016-2017. 

```{r closure, message = FALSE}
closure_df = 
  rest_inspec %>% 
  filter(str_detect(action, "[cC]losed") == TRUE,
    critical_flag == "Critical",
    inspection_date > "2016-01-01" & inspection_date <= "2017-12-31"
  ) %>% 
  select(-c(building, street, phone, record_date)) %>% 
  mutate(
    inspection_date = as.Date(inspection_date),
    camis = as.character(camis),
    dba = str_to_title(dba)
  )

closure_plot = 
  closure_df %>%
  mutate(
    text_label = str_c(
      "", dba,
      "\n", inspection_date,
      "\nScore: ", score
    )
  ) %>% 
  plot_ly(x = ~inspection_date, y = ~score, color = ~boro, type = "scatter", 
    colors = "viridis", alpha = 0.5, text = ~text_label) %>% 
    layout(xaxis = list(title = "Inspection Date"), 
         yaxis = list(title = "Score"))

closure_plot  
```

While it can be a bit difficult to tell, this plot shows that there were more inspections that resulted in closure in 2017 overall, compared to 2016. There were also two standout restaurants: Margarita Island in Brooklyn, which had a score of 151, and Maman in Manhattan, with a score of 145. The darker a circle is, the more violations that restaurant had at that inspection that ultimately caused it to close. Number of violations is directly related to score, which is why the higher scores have darker points, and lower scores have lighter points. There does appear to be one likely entry error, with I Love NY Pizza receiving a score of 0.

### Dashboard

To see the composite dashboard of this analysis, please visit the [dashboard](dashboard.html) page. 

